---
title: "High Dimensional Exercise 4"
author: 'Hamed Vaheb'
output:
  pdf_document: default
  html_document: default
date: '2022-03-22'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(dplyr)
library(ggplot2)
library(broom)
```

In an experiment on the analysis of the link between the beats per minute under stress and the age of a sample of 10 men, the following data have been collected:
\begin{table}[h]
        \centering
        \begin{tabular}{||c|c|c|c|c|c|c|c|c|c|c||}
        \hline
{\textbf{Beats per minute}}  & 200 & 195 & 200 & 190 & 188 & 180 & 185 & 180 & 163 & 170 \\ \hline
{\textbf{Age}} & 10 & 20 & 21 & 25 & 29 & 30 & 31 & 40 & 45 & 50 \\ \hline
        \end{tabular}
\end{table}
\begin{enumerate}
    \item Draw  the scatterplot.
    \item Build the linear regression model  $Y_i = b_0 + b_1 t_i + \epsilon_i$ and estimate $b_0$ and $b_1$.
    \item Verify the null hypothesis $H_0 : b_1 = 0$ against the alternative $H_1: b_1 \neq 0$ with a significance level $\alpha = 0.05$. 
\end{enumerate}


# **Prepare**
First we insert data of Beats per minute and age as columns of a new dataframe.

```{r cars}
BeatsPerMin <- c(200, 195, 200, 190, 188, 180, 185, 180, 163, 170)
Age<- c(10, 20, 21, 25, 29, 30, 31, 40, 45, 50)

df <- data.frame(BeatsPerMin, Age)



```

## Question 1: Drawing Scatterplot
Draw the scatterplot
```{r pressure}
x <- Age
y <- BeatsPerMin
# Plot with main and axis titles
# Change point shape (pch = 19) and remove frame.
# Add regression line
plot(x, y, main = "Scatter plot",
     xlab = "age", ylab = "Beats per minute",
     pch = 19, frame = FALSE)
abline(lm(y ~ x, data = df), col = "blue")
```

## Question 2: Regression Model
Build the linear regression model $Y_i = b_0 + b_1 \cdot t_i + \epsilon_i$ and estimate $b_0$ and $b_1$.

$b_0 = 213.1721, b_1 = -0.9326$ 

```{r}
model <- lm(BeatsPerMin ~ Age, data = df)
summary(model)
```

## Question 3: Hypothesis Testing
Verify the null hypothesis $H_0 : b_1 = 0$ against the alternative $H_1 : b_1 \neq 0$ with a significance level $\alpha = 0.05$.

The lm function provides p-values for significance of both intercept and slope. Since we are interested in the slope, we can compare the p-value (0.000101) with $\alpha = 0.05$, concluding that p-value is less than alpha, we reject the null hypothesis. This underlines that the dependent variable (age) has influence of target variable (beats per minute).

