---
title: "High Diminsional Statistics-Sheet 2-Exercise 4"
author: "Hamed Vaheb"
date: "3/31/2022"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#library(tidyverse)
library(knitr)
library(dplyr)
library(ggplot2)
library(broom)
library(reshape2)
#library(ggpubr)
```

The goal of an experimental research is to understand whether or not some hormonal treatments
have effects on the hormonal concentrations of 12 adult female dogs.
No treatment has been given to 4 dogs. Their hormonal concentrations are 117, 124, 40 and 88.
4 dogs have been treated with oestrogens. Their hormonal concentrations after the treatment
are 440, 264, 221, 136.
4 dogs have been treated with progesterone. Their hormonal concentrations after the treatment
are 605, 626, 385, 475.

## Prepare

```{r cars}
index <- c(1:4)
cnctr_before <- c(117, 124, 40, 88)
cnctr_after_oest <- c(440, 264, 221, 13)
cnctr_after_progest <- c(605, 626, 385, 475)
```

### Question 1: Plotting Concentration by Treatment
Plot a multiple histogram showing hormonal concentrations by treatment.

#### Scatterplot

```{r pressure}
# Plot with main and axis titles
# Change point shape (pch = 19) and remove frame.
# Add regression line
df <- data.frame(index, cnctr_before, cnctr_after_oest, cnctr_after_progest)

df_plot <- melt(df, id.vars = 'index', variable.name = 'series')


ggplot(data = df_plot, aes(index, value)) +
  geom_point(aes(colour = series)) +
  labs(x = "Treatment", y = "Hormonal Concentration")


```

#### Histogram


```{r}

set.seed(1)
df
df_plot <- df_plot[order(df_plot$index),]

df_plot 
df_plot$series <- as.factor(df_plot$series)


ggplot(df_plot, aes(x=value, fill=series)) +
  geom_histogram( color='#e9ecef', alpha=0.6, position='identity')
```

### Question 2: Table of man and stds
Calculate a table that shows the mean and standard deviation of hormonal concentrations by treatment.

```{r S}

 sd <- df %>% select(c('cnctr_before', 'cnctr_after_oest', 'cnctr_after_progest')) %>% apply(2,sd)

mean <- df %>% select(c('cnctr_before', 'cnctr_after_oest', 'cnctr_after_progest')) %>% apply(2,mean)

rbind(mean, sd)

```

### Question 3: ANOVA
Use ANOVA to test for a difference between treatments in the hormonal concentrations
of the dogs, with a significance level $/alpha = 0.05$. What is your conclusion about the effect
of the treatments?

One-way analysis of variance (ANOVA) is employed when there exists a categorical independent variable (with two or more categories) and a normally distributed interval dependent variable and we aim at testing the difference between means of the dependent variable on broken down by the levels of the independent variable.

We assume that the categorical independent variable is hormonal treatment, which consists of two groups, namely oestrogens and progesterone. Mean of each group is denoted by $\mu_O$ and $\mu_P$ respectively. We also assume the dependent variable is hormonal concentrations. We formulate our hypothesis test as the following: 

\(H_{0}:\mu_{O} - \mu_{P} = 0\) \hspace{0.5cm} against \hspace{0.5cm} \(H_{1}:\mu_{O} - \mu_{P} \neq 0\)

```{r Ss}

aov.model <- aov(df_plot$value ~ df_plot$series)
summary(aov.model)
aov.model$coefficients
```


The associated p-value helps to interpret the F-statistic. Given significance level $\alpha =0.05$, since p_value = 0.00234 and hence p_value <$\alpha$, we reject null hypothesis and we increase our certainty in $H_1$, which states that there is significant difference between mean of concentration level of the groups that has been treated with oestrogen and the the one that has been treated with progesterone.
 